---
title: "Writing reports with R markdown"
author: "Esmée Kuiper"
date: "2022-12-09"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 987)
```

```{r packages, include=FALSE}
library(tidyverse)
library(here)
library(knitr) #for kable
```

This report was prepared to the attention of the UN. It analyzes the relationship between a country's GDP, its life expectancy, and CO2 emissions. Our goal is to determine what to degree a country's economic strength or weakness is determined by its public health and climate status. We hypothesize that both life expectancy and CO2 emissions increase with a country's GDP. 

```{r data, include=FALSE}
gapminder_1997 <- read_csv(here("data/gapminder_1997.csv"))

n_countries <- gapminder_1997 %>% 
  select(country) %>% 
  n_distinct()

min_gdp <- gapminder_1997 %>% 
  summarize(min=min(gdpPercap)) %>% 
  pull(min) %>% 
  round() 

max_gdp <- gapminder_1997 %>% 
  summarize(max=max(gdpPercap)) %>% 
  pull(max) %>% 
  round()
```

```{r gdp_lifeExp_1997,echo=FALSE}
gapminder_1997 %>% 
  ggplot(aes(x=gdpPercap, y=lifeExp, color=continent, size=pop/1e6)) +
  geom_point() +
  labs(x = "GDP per capita", y = "life expectancy", 
       title = "Do people in wealthy countries live longer?",
       size = "Population (in millions)")
```

The above plot shows the relationship between GDP per capita and life expectancy for a total of `r n_countries` countries. For this set of countries, economic wealth ranged from a minimum of USD `r min_gdp` to a maximum of USD `r max_gdp` per capita. 

## Some type of title

We can make tables in R markdown or by using the kable function in the knitter package: 

| Summary of data |
|----------|----------|
| Number of countries | `r n_countries`  |
| Maximum GDP (USD) | `r min_gdp` |
| Minimum GDP (USD) | `r max_gdp` |

```{r oceania_table, echo=FALSE}
gapminder_1997 %>% 
  filter(continent == "Oceania") %>% 
  kable() # if you want more power: kableExtra
```

#### Practice problems

1. **R**  
2. *ggplot*  
3. ***more on data cleaning and analysis***  
4. [Cheat sheet](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)


