---
title: "Writing reports with R markdown"
author: "Esmée Kuiper"
date: "2022-12-09"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(here)
```

This report was prepared to the attention of the UN. It analyzes the relationship between a country's GDP, its life expectancy, and CO2 emissions. Our goal is to determine what to degree a country's economic strength or weakness is determined by its public health and climate status. We hypothesize that both life expectancy and CO2 emissions increase with a country's GDP. 

```{r data, include=FALSE}
gapminder_1997 <- read_csv(here("data/gapminder_1997.csv"))
```

```{r gdp_lifeExp_1997,echo=FALSE}
gapminder_1997 %>% 
  ggplot(aes(x=gdpPercap, y=lifeExp, color=continent, size=pop/1e6)) +
  geom_point() +
  labs(x = "GDP per capita", y = "life expectancy", 
       title = "Do people in wealthy countries live longer?",
       size = "Population (in millions)")
```


